Supplementary Material

### 2013-10-04 icml/lower-bounds.tex

Although it is possible to achieve the fast $O(1/T^{2})$ convergence rate for finite-dimensional herding, \citet{Bach:2012a} note that without further assumptions herding has only a $O(1/T)$ convergence rate in the infinite-dimensional case. We now provide an example showing rate is tight for {\em any} herding-like algorithm.  By ``herding-like,'' we mean algorithms that approximate the distribution with at most one additional sample at each step. We formalize this idea by defining a {\em sparse update method} to be any method that produces an estimator of $\bar \phi$ at iteration $t$ of the form 
\[
\theta_{t} = \sum_{s =1}^{t}w_{t} \phi(x_{t}).
\]
Sparse update methods include $\alpha_{t}$-weighted herding and \cgd with line search. These sparse methods -- which only operate on the extreme points of the simplex -- are of interest because they provide (weighted) ``samples'' matching the moments of the distribution of interest.

Our example is as follows. Let $\sX$ be the set of vectors $(x_{1},x_{2},\dots) \in [0,1]^{\infty}$ such that $\sum_{n=1}^{\infty} x_{n} = 1$ and a finite number of the $x_{n}$ are non-zero. Define the infinite feature vector $\phi$ to be the imbedding of $\sX$ into $\ell^{1}$. Let the target distribution $p(x)$ be chosen such that $\bar\phi_{n} = c/\alpha(n)$, where $\alpha(n)$ is any superlinear function and $c^{-1} = \sum_{n=1}^{\infty} \alpha(n) < \infty$. Without loss of generality we can assume that $\alpha$ is monotonically increasing. We will consider the standard $\ell^{2}$ herding loss function $L(\theta) = \|\theta - \bar \phi\|_2^2$.

Since we are only considering sparse update methods, the $t$-th estimator will have at most $t$ non-zero entries, and so 
\[
\|x-x^{*}\|_2^2 \ge c^{2}\sum_{n > t} 1/\alpha(n)^{2} = \Theta(1/\beta(t)) \label{eq:convergence-lb}
\]
for some (monotonically increasing) superlinear function $\beta$. But for any such $\beta$ we can choose some $\alpha$ to obtain \eqref{eq:convergence-lb}. In particular, let $1/\gamma(t) = \frac{\dee}{\dee t}\frac{1}{\beta(t)}$, where $\gamma(t) \in \omega(t^{2})$. Then we can choose $\alpha = \sqrt \gamma \in \omega(t)$ to satisfy \eqref{eq:convergence-lb}.

Therefore, any sparse update method must have convergence rate no faster than $\Theta(1/\beta(t))$ for all superlinear $\beta$, and hence no faster than $\Theta(1/t)$.

>>>>

Although it is possible to achieve the fast $O(1/T^{2})$ convergence rate for finite-dimensional herding, \citet{Bach:2012a} note that without further assumptions herding has only a $O(1/T)$ convergence rate in the infinite-dimensional case. We now provide an example showing this bound is tight for {\em any} herding-like algorithm.  By ``herding-like,'' we mean algorithms that approximate the distribution with at most one additional sample at each step. We formalize this idea by defining a {\em sparse update method} to be any method that produces an estimator of $\bar \phi$ at iteration $t$ of the form 
\[
\theta_{t} = \sum_{s =1}^{t}w_{t} \phi(x_{t}),
\]
where $w_{t} \in \bR$. Sparse update methods include herding, $\alpha_{t}$-weighted herding, \cgd, and  \cgd with line search. These sparse methods -- which only operate on the extreme points of the simplex -- are of interest because they provide (weighted) ``samples'' matching the moments of the distribution of interest.

Our example is constructed as follows. Let the sample space be $\sX = \{ 1, 2, 3, \dots\}$. Define the infinite feature vector $\phi$ to be $\phi_{n}(x) = \bI(x = n)$, so $\sM$ is the unit $\ell^{1}$ ball. For target distribution $p \in \Delta_{\sX}$, we therefore have  that the mean vector is the vector of probabilities: $\bar\phi_{n} = p(n)$. Let $p(x) = c/\alpha(n)$, where $\alpha(n)$ is any superlinear function (i.e.~$\alpha \in \omega(n)$) and $c^{-1} = \sum_{n=1}^{\infty} \alpha(n) < \infty$. Without loss of generality we can assume that $\alpha$ is monotonically increasing.  We will consider the standard $\ell^{2}$ herding loss function $L(\theta) = \|\theta - \bar \phi\|_2^2$.

Since we are only considering sparse update methods, the $t$-th estimator will have at most $t$ non-zero entries, and so 
\[
\|\theta_{t}- \bar \phi\|_2^2 \ge c^{2}\sum_{n > t} 1/\alpha(n)^{2} = \Theta(1/\beta(t)) \label{eq:convergence-lb}
\]
for some (monotonically increasing) superlinear function $\beta$. But for any such $\beta$ we can choose some $\alpha$ to obtain \eqref{eq:convergence-lb}. In particular, let $1/\gamma(t) = \frac{\dee}{\dee t}\frac{1}{\beta(t)}$, where $\gamma(t) \in \omega(t^{2})$. Then we can choose $\alpha = \sqrt \gamma \in \omega(t)$ to satisfy \eqref{eq:convergence-lb}. Therefore, any sparse update method must have convergence rate no faster than $\Theta(1/\beta(t))$ for all superlinear $\beta$, and hence no faster than $\Theta(1/t)$.

### 2013-10-04 icml/supplementary-material.tex
draft of connections subsection

The \bmd algorithms are very closely related to \cgd and \md. \dual and \cgd are related as follows. \dual minimizes $g + R^{*}$ while \cgd minimizes $h^{*} + f^{*}$. Make the identifications $h^{*} = R^{*}$ and $f^{*} = g$. Note that $R^{*}$ is $\beta$-smooth $\iff R$ is $(1/\beta)$-strongly convex $\iff h$ is $(1/\beta)$-strongly convex, so the conditions on the functions being optimized are equivalent. Then with $\rho_{t} = \alpha_{t+1}/A_{t+1}$, we can rewrite the \cgd updates  s 
\[
x_{t} &= \partial R^{*}(-\bar y_{t})  \\
y_{t} &= \partial g^{*}(x_{t}) \\
\bar y_{t+1} &= A_{t+1}^{-1}\sum_{s \le t+1} \alpha_{t} y_{t},
\]
which is identical to the \dual if we make the substitutions $x_{t} \to -\theta_{t}$ and $y_{t} \to -u_{t}$.

We similarly relate \primal and \md. \primal minimizes $g + R^{*}$ while \md minimizes $h + f$. Make the identifications $h = g$ and $f = R^{*}$ and note the conditions of $h$ and $g$ are the same. Letting $\rho_{t+1} = \alpha_{t}/A_{t}$, we can rewrite the \md updates as 
\[
y_{t} &= \partial R^{*}(x_{t}) \\
x_{t+1} &= \partial g^{*}((1 - \rho_{t+1})\partial g(x_{t}) - \rho_{t+1}y_{t}) \\
&= \partial g^{*}(-\bar y_{t}), 
\]
where $\bar y_{t}$ is the same as above. Making the substitutions $y_{t} \to \theta_{t}$ and $x_{t} \to u_{t}$ completes the identification of the two methods.

>>>>

The \bmd algorithms are very closely related to \cgd and \md. \dual and \cgd are related as follows. \dual minimizes $g + R^{*}$ while \cgd minimizes $h^{*} + f^{*}$. Make the identifications $h^{*} = R^{*}$ and $f^{*} = g$ and note the conditions on the functions being optimized are equivalent. Then with $\rho_{t} = \alpha_{t+1}/A_{t+1}$, we can rewrite the \cgd updates as 
\[
x_{t} &= \partial R^{*}(-\bar y_{t})  \\
y_{t} &= \partial g^{*}(x_{t}) \\
\bar y_{t+1} &= A_{t+1}^{-1}\sum_{s \le t+1} \alpha_{t} y_{t},
\]
which is identical to the \dual if we make the substitutions $x_{t} \to -\theta_{t}$ and $y_{t} \to -u_{t}$.

We similarly relate \primal and \md. \primal minimizes $g + R^{*}$ while \md minimizes $h + f$. Make the identifications $h = g$ and $f = R^{*}$ and note the conditions of $h$ and $g$ are the same. Letting $\rho_{t+1} = \alpha_{t}/A_{t}$, we can rewrite the \md updates as 
\[
y_{t} &= \partial R^{*}(x_{t}) \\
\bar y_{t} &= \partial g^{*}((1 - \rho_{t+1})\partial g(x_{t}) - \rho_{t+1}y_{t}) \\
x_{t+1} &= \partial g^{*}(-\bar y_{t}), 
\]
where $\bar y_{t}$ is the same as above. Making the substitutions $y_{t} \to \theta_{t}$ and $x_{t} \to u_{t}$ completes the identification of the two methods.

Finally, note that if we write $\alpha_{t} = \rho_{t}/\prod_{s=1}^{t} (1 - \rho_{s})$, we have $A_{t} = \prod_{s=1}^{t} (1 - \rho_{s})$ and hence $\alpha_{t} / A_{t} = \rho_{t}$. Hence, up to a change in index, we have a bijection between the $\rho_{t}$ and the $\alpha_{t}$.

\section{Proof Lemma used in $q$-Herding Convergence Proof}

Two simple equalities were used in the proof of the convergence rate of $q$-herding. We now provide proofs. Using the fact that $x^{n} - y^{n} = (x - y)\sum_{i=0}^{n-1}x^{i-1}y^{n-i-1}$, for the first inequality we have
\(
x^{p} - py^{p-1}x + (p-1)y^{p}
&= x^{p} - y^{p} - py^{p-1}x + py^{p} \\
&= (x - y)\sum_{r=0}^{p-2} (y^{r}x^{p-r-1} - y^{p-1}) \\
&= (x - y)^{2}\sum_{r=0}^{p-2}\sum_{s=0}^{p-r-2}x^{p-r-s-2}y^{r+s} \\
&\le (x - y)^{2}\sum_{r=0}^{p-2}\sum_{s=0}^{p-r-2}\max(x, y)^{p-2} \\
&= {p \choose 2}(x - y)^{2}\max(x, y)^{p-2}.
\)
For the second inequality we have
\(
(y-z)^{2}\max(x,y)^{p-2} 
&\le 4\max(y,z)^{2}\max(x,y)^{p-2} \\
&\le 4\max(x,y,z)^{p}.
\)
